---
title: "Quiz 3 Getting and Cleaning Data Quiz 3"
author: "Cesar Espitia"
date: "June 16, 2015"
output: html_document
---

Question 1
Create a logical vector that identifies the households on greater than 10 acres who sold more than $10,000 worth of agriculture products. Assign that logical vector to the variable agricultureLogical. Apply the which() function like this to identify the rows of the data frame where the logical vector is TRUE. which(agricultureLogical) What are the first 3 values that result?


```{r}
library(dplyr)
data <- read.csv("getdata-data-ss06hid.csv", header=T)
agricultureLogical <- filter(data, ACR == 3, AGS ==6)
which(agricultureLogical)[1:3]
```

Question 2

```{r, echo=FALSE}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fjeff.jpg"
> file <- file.path(getwd(), "jeff.jpg")
> download.file(url, file, mode = "wb", method = "curl")
> img <- readJPEG(file, native = TRUE)
> quantile(img, probs = c(0.3, 0.8))
```

Question 3:

```{r, echo=FALSE}
> url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv"
> file <- file.path(getwd(), "GDP.csv")
> download.file(url, file, method = "curl")
> dtGDP <- data.table(read.csv(file, skip = 4, nrows = 215))
> dtGDP <- dtGDP[X != ""]
> dtGDP <- dtGDP[, list(X, X.1, X.3, X.4)]
> setnames(dtGDP, c("X", "X.1", "X.3", "X.4"), c("CountryCode", "rankingGDP", "Long.Name", "gdp"))
> url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
> file <- file.path(getwd(), "EDSTATS_Country.csv")
> download.file(url, file, method = "curl")
> dtEd <- data.table(read.csv(file))
> dt <- merge(dtGDP, dtEd, all = TRUE, by = c("CountryCode"))
> sum(!is.na(unique(dt$rankingGDP)))
[1] 189
> dt[order(rankingGDP, decreasing = TRUE), list(CountryCode, Long.Name.x, Long.Name.y, rankingGDP, gdp)][13]
   CountryCode         Long.Name.x         Long.Name.y rankingGDP   gdp
1:         KNA St. Kitts and Nevis St. Kitts and Nevis        178  767 
```

```{r, echo=FALSE}
> dt[, mean(rankingGDP, na.rm = TRUE), by = Income.Group]
           Income.Group        V1
```

```{r, echo=FALSE}
> breaks <- quantile(dt$rankingGDP, probs = seq(0, 1, 0.2), na.rm = TRUE)
> dt$quantileGDP <- cut(dt$rankingGDP, breaks = breaks)
> dt[Income.Group == "Lower middle income", .N, by = c("Income.Group", "quantileGDP")]
```

```{r, echo=FALSE}
```

```{r, echo=FALSE}
```